Towards Spherical Robots For Mobile Mapping In Human Made Environments\tnoteref{mytitlenote}

### 2021-07-07 Latex/rebuttal.txt

> - Section 3.2 deals with the extraction of planes, but the following
> questions remain unclear: Are the planes always searched for only in
> the context of the observational data of the current epoch or with
> regard to the already registered global point clouds? How many
> planes are needed at all and what demands are made on their spatial
> distribution? The abstract mentioned parallel and perpendicular
> planes, but this is not discussed further here. Are only these used
> and if so, why? Furthermore, what about planes that are not
> necessarily parallel or perpendicular (e.g. doors, cupboards)? How
> is the flatness of a plane checked? How are outliers dealt with,
> especially with regard to existing measurement and motion noise?

We added more detail in the plane extraction subsection.

> - Throughout Section 3, numerous variables are not introduced and
> are not explained in more detail, so that a more precise
> interpretation of the individual equations is unfortunately very
> difficult.

TODO: Anton? Have you done it already? 
Yes - Anton

> 
> - What orders of magnitude were finally used for the range noise
> model in Section 4.2 and why was this not also done for the angular
> resolution?

The distributions we sample from are specified in the caption of 
figure 3.

> - In the context of the real experiment in Section 5, a major
> simplification is made. Even if this simplification is certainly
> suitable for the beginning, this special case comes quite surprising
> here...especially in comparison to the simulation and the previously
> described topic. Have other measurements not yet been carried out or
> what is the background? At least this simplification should be
> mentioned beforehand, otherwise the reader will have false
> expectations.

We added a further experiment that uses nearly all DoF: A sphere rolling
on a ground plane. This was not done in the previous manuscript because
we had issues getting the hardware shipped, assembled and the experiment 
carried out in time.

> - Wouldn't it also be interesting to compare the estimated pose
> parameters with the nominal values from the simulation within the
> framework of Section 6.1? So far, the focus has only been on the
> point cloud.

FIXME: Do we have space & time for this?

> - Why is there no qualitative validation for the results of the real
> data in Section 6.2? A reference scan based on a terrestrial laser
> scanner would be a possibility or otherwise based on the flatness of
> known planes in the registered point cloud?

FIXME: Can we get a riegl scan of jaspers home? If not corona restrictions were at fault
We couldn't access the rooms the dataset was acquired in for a terrestrial 
reference scan.

> - In the Conclusion, the parallelism and perpendicularity within the
> environment is mentioned again, although this did not play a role in
> the main part of the paper. Therefore, the question arises why do
> the planes have to be parallel or perpendicular at all and why is
> this mentioned at the beginning? This is not clear, since no
> additional constraints on this additional information have been
> introduced so far. There are also already other methods available
> that take this idea of geometric constraints for georeferencing into
> account.

We removed the initial mention of parrallelism and perpendicularity of the planes
as it is of no further interest in the rest of the manuscript.

> Further open questions/comments are:
> Page 1, Line 15: Double use of "protect", which should be replaced
> by another word.

### 2021-07-07 Latex/sections/conclusion.tex
LPC results added, rolling sphere added.
Proofread & Added some answers to rebuttall.txt

In this paper, we proposed an approach to mobile mapping using a spherical robot. 
Using the inherent rotation of the sphere used for locomotion to also rotate a laser scanner also for measurements of the entire environment.
We propose to post-process the acquire data using a novel registration method for man made environments that exploits the structure of those environments. 
Given that in human-made environments straight planes are abundantly available, we employ point-to-plane correspondences to improve a pre-registered 3D point cloud. 
We have evaluated the procedure on a simulated dataset and on an experimentally acquired dataset with different motion profiles. 
In this evaluation, we have shown that the procedure improves both datasets and yields maps that better resemble human-made, structures. 
In particular, the qualitative structure of the environments is reconstructed well. 
In the resulting maps, the parallel walls are clearly improved.

@@

%
The biggest issue with the presented algorithm currently is that the global plane model is established only once as initialization, but never updated afterwards.
Utilizing the local planar clustering (LPC) of the individual line scans for updating the global plane model step by step is currently the primary objective of future studies.
One idea is to calculate the convex hulls of the LPC, and combine them with the global planes sequentially.
Therefore, we expect to increase robustness and autonomy of the registration procedure.

%
Currently, we are assembling the sphere of Figure~\ref{fig:robotRender} to test it in real environments.
Unfortunately, this sphere will be too large to fit our floating setup, thus it will be tested in an office like environment.
Furthermore, we will test our rolling and scanning spheres on a rail system to achieve repeatable trajectories.

>>>>

In this paper, we proposed an approach to mobile mapping using a spherical robot. 
Given that a spherical robot inherently rotates for locomotion, we use this movement to also rotate a laser scanner, that consequently measures the robots entire environment.
We propose to post-process the acquire data using a novel registration method for man made environments that exploits the structure of those environments. 
In human-made environments straight planes are abundantly available, hence we employ point-to-plane correspondences to improve a pre-registered 3D point cloud. 
We have evaluated the procedure on a simulated dataset and on two experimentally acquired datasets with different laser scanners and motion profiles. 
In this evaluation, we have shown that the procedure improves all datasets and yields maps that better resemble human-made, structures. 
In particular, the qualitative structure of the environments is reconstructed well. 
In the resulting maps, the parallel walls are clearly improved.

@@

%
The biggest issue with the presented algorithm currently is that the global plane model is established only once at initialization, but never updated afterwards.
Utilizing the local planar clustering (LPC) of the individual line scans for updating the global plane model step by step is currently the primary objective of future studies.
One idea is to calculate the convex hulls of the LPC, and combine them with the global planes sequentially.
Therefore, we expect to increase robustness and autonomy of the registration procedure.

%
Furthermore, more experimental evaluation is required, hence we will test our rolling and scanning spheres on a rail system to achieve repeatable trajectories.

### 2021-07-07 Latex/sections/realData.tex
Proofread & Added some answers to rebuttall.txt

The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor that the simulation uses.
We put the laserscanner into an acrylic glass spherical shell together with three IMUs and roll it by hand with a slow, constant motion in a home indoor environment.
Thus, this time almost all six degrees of freedom are used for optimization.
We exlude a translation in the global up- or downwards direction, though, since we also locked that dimension when collecting the data.
Figure~\ref{fig:prototype} shows the prototpye.
The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the pure transformation found by the recorded pose.
Although the robot has been moved carefully slow, you see that pose error accumulates in such a way that the same wall is visible multiple times.

\begin{figure}
  \includegraphics[width=0.49\linewidth]{./images/photoReal1.jpeg}
  \includegraphics[width=0.50\linewidth]{./images/photoReal3.jpeg}
  \caption{Prototype used for data acquisition with the rolling sphere. The main payload is the Livox Mid-100 laser-scanner. For pose-estimation, three IMUs of the manufacturer Phidget are placed inside and a Raspberry Pi 4 for the calculations. On the top are two batteries, and on the bottom one voltage stabilizer and breakout box of the laser-scanner. }
  \label{fig:prototype}
\end{figure}

>>>>

The second datasets is collected by a 3D scanner, namely the LIVOX MID 100, which is the same sensor the simulation uses.
We put the laser scanner into an acrylic glass spherical shell together with three IMUs  (same specifications as in the first dataset) and manually roll it with a slow, constant motion in a home indoor environment.
Thus, this time nearly all six degrees of freedom are used for optimization.
We exclude a translation in the global up- or downwards direction as the sphere rolls on the ground surface, without possibility for up- or downwards motion-
Figure~\ref{fig:prototype} shows the prototype.
The left column of figure~\ref{fig:jasperhome} shows the pre-registered point cloud, applying only the transformation defined by the recorded poses.
Although the robot has been moved carefully, you see that pose error accumulates in such a way that the same wall is sensed multiple times.

\begin{figure}
  \includegraphics[width=0.49\linewidth]{./images/photoReal1.jpeg}
  \includegraphics[width=0.50\linewidth]{./images/photoReal3.jpeg}
  \caption{Prototype used for data acquisition with the rolling sphere. The main payload is the Livox Mid-100 laser scanner. For pose-estimation, three IMUs of the manufacturer Phidget are placed inside and a Raspberry Pi 4 for the calculations. On the top are two batteries, and on the bottom one voltage stabilizer and the breakout box of the laser scanner. }
  \label{fig:prototype}
\end{figure}

### 2021-07-07 Latex/sections/registration.tex
Proofread & Added some answers to rebuttall.txt

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as the square of the L2-norm of the distance:
\begin{align}
	E = \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_k} \sum_{\vec{p}_i\in \mathcal{P}_k}2D(\vec{\Pi}, \vec{p}_i)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{equation}
\begin{split}
\nabla E_\varphi = n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]+y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] - z_i C_\varphi C_\vartheta)\\
        + n_{\mathcal{P}_k}^z(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] +y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
\end{split} \;\; ,
\end{equation}
\begin{equation}
\begin{split}
 \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
        &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi 
\end{split}
\;\; ,
\end{equation} 
\begin{equation}
\begin{split}
	\nabla E_\psi = &n_{\mathcal{P}_k}^x\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
       &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi] 
       + y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
       &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] 
       + y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
\end{split}
\end{equation}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

@@

Firstly, we introduce the possibility to choose a lock on some dimensions from being optimized by setting the corresponding $\alpha_i$ to zero.
Although 6D optimization generally works, reducing the optimization space is particularly useful if the source of error in the system is known and a model exists.
That way, e.g., as we expect a spherical robot to move on a plane the position along the axis perpendicular to the plane is constant and should not be used for optimization.\todo{@Fabi clarify in experiments which locks were active.}

>>>>

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as the square of the L2-norm of the distance:
\begin{align}
	E = \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_k} \sum_{\vec{p}_i\in \mathcal{P}_k}2D(\vec{\Pi}, \vec{p}_i)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{equation}
\begin{split}
\nabla E_\varphi = &n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]\\
			       & +y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] - z_i C_\varphi C_\vartheta)\\
                   &+ n_{\mathcal{P}_k}^z(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\
                   &+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
\end{split} \;\; ,
\end{equation}
\begin{equation}
\begin{split}
 \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
        &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi 
\end{split}
\;\; ,
\end{equation} 
\begin{equation}
\begin{split}
	\nabla E_\psi = &n_{\mathcal{P}_k}^x\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
       &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi]\\
       &+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
       &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
       &+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
\end{split}
\end{equation}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

@@

Firstly, we introduce the possibility to choose a lock on some dimensions from being optimized by setting the corresponding $\alpha_i$ to zero.
Although 6D optimization generally works, reducing the optimization space is particularly useful if the source of error in the system is known and a model exists.
That way, e.g., as we expect a spherical robot to move on a plane the position along the axis perpendicular to the plane is constant and should not be used for optimization.

### 2021-07-07 Latex/sections/results.tex
Proofread & Added some answers to rebuttall.txt

Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
The pre-registration is obtained by determining the orientation of the sphere via Madgwick filtered~\cite{madgwick2010efficient} IMU measurements.
To increase the number of possible point-to-plane correspondences we combine twenty temporally successive scans into one meta-scan, which is then globally registered. 
We always use the middlest line scan as a reference coordinate system.
This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming the scan in a senseless way.
Transformations like these happen in particular for a small collection of points as outliers have more influence.

@@

Figure~\ref{fig:jasperhome} shows the results obtained before and after employing the presented algorithm on the dataset, acquired by the prototype in figure~\ref{fig:prototype}.
We again combine twenty temporally successive scans into one meta-scan, using the middlest scan as a reference coordinate system, which is then globally registered.

\begin{figure*}
        \centering
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_corner}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_corner}\\
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_top}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_top}
        \caption{The point cloud acquired by the rolling sphere before (left) and after (right) applying the plane based registration. View from the interior (top) and a birds-eye view (bottom). Parameters used for optimization: $\epsilon_H = 200$, $\epsilon_P = 300$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 200$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 800$, $\alpha_0 = [0.001, 0.001, 0.001, 1, 0, 1]^\tau$, $i = 1500$, $k = 1$. An animation of the registration process is given at \url{https://youtu.be/hYR3xLhIdsE}.}
        \label{fig:jasperhome}
\end{figure*}
After the registration, the walls of the environment are no longer ambiguous, since all the sub-sequent scans got matched with the initial plane representation.
However, the improved path looks distorted.
This is because for some scans, especially empty ones where the sensor points to the ground, no planar features got matched with the global plane model, thus their pose is not optimized.
Improving the path quality is an objective for future works.

>>>>

Figure~\ref{fig:cylon-corrected} shows the results obtained before and after employing the plane based registration on the dataset acquired by the floating sphere.
The pre-registration is obtained by determining the orientation of the sphere via Madgwick filtered~\cite{madgwick2010efficient} IMU measurements.
To increase the number of possible point-to-plane correspondences we combine twenty temporally successive scans into one meta-scan, which is then globally registered. 
We always use the scan at the median index as a reference coordinate system.
This does not only speed up convergence due to the proportional effect on the error function but also decreases the risk of transforming a single line scan incorrectly.
Transformations like these happen in particular for a small collection of points as outliers have more influence.

@@

Figure~\ref{fig:jasperhome} shows the results obtained before and after employing the presented algorithm on the dataset, acquired by the prototype in figure~\ref{fig:prototype}.
We again combine twenty temporally successive scans into one meta-scan, using the scan at the median index as a reference coordinate system, which is then globally registered.

\begin{figure*}
        \centering
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_corner}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_corner}\\
        \includegraphics[width=0.495\textwidth]{./images/jasper_uncorr_top}\hfill
        \includegraphics[width=0.495\textwidth]{./images/jasper_corr_top}
        \caption{The point cloud acquired by the rolling sphere before (left) and after (right) applying the plane based registration. View from the interior (top) and a birds-eye view (bottom). Parameters used for optimization: $\epsilon_H = 200$, $\epsilon_P = 300$, $\epsilon_\alpha = \frac{\pi}{4}$, $K = 200$ for AKNN, $d_{growth} = 50$, $N_{c_min} = 800$, $\alpha_0 = [0.001, 0.001, 0.001, 1, 0, 1]^\tau$, $i = 1500$, $k = 1$. An animation of the registration process is given at \url{https://youtu.be/hYR3xLhIdsE}.}
        \label{fig:jasperhome}
\end{figure*}
After the registration, the walls of the environment are no longer ambiguous, since all the subsequent scans are matched with the initial plane representation.
When considering the resulting path, it looks distorted.
This is because for some scans, especially empty ones where the sensor points to the ground, no planar features got matched with the global plane model, thus their pose is not optimized.
Improving the path quality is an objective for future works.

### 2021-07-07 Latex/sections/simulation.tex
Proofread & Added some answers to rebuttall.txt

The Livox-Mid100 laser scanner consists of three Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (horizontal) by \SI{38.4}{\degree} (vertical).
It provides \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range~\cite{LivoxMid40-100}.
A particular advantage of the Livox laser scanners is their unique scanning pattern.
The scanners use two motorized, so-called Risley prisms to steer the beam deflectance.
For the specific ratio of rotation speeds used in the Livox scanners, the beam traces out a non-repetitive, flower-shaped scanning pattern~\cite{thorlabs}.
Since the pattern is non-repetitive, the point density increases with integration time, enabling denser measurements.

>>>>

The Livox-Mid100 laser scanner consists of three Mid40 scanners arranged to provide an overall field-of-view of \SI{98.4}{\degree} (horizontal) by \SI{38.4}{\degree} (vertical).
It provides a rate of \SI[per-mode=symbol]{300,000}{\pts\per\second} with a range precision of \SI{2}{\centi\meter} up to \SI{20}{\meter} range~\cite{LivoxMid40-100}.
A particular advantage of the Livox laser scanners is their unique scanning pattern.
The scanners use two motorized, so-called Risley prisms to steer the beam deflectance.
For the specific ratio of rotation speeds used in the Livox scanners, the beam traces out a non-repetitive, flower-shaped scanning pattern~\cite{thorlabs}.
Since the pattern is non-repetitive, the point density increases with integration time, enabling denser measurements.
